\documentclass[11pt,letterpaper]{article}
%%%%%\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{array}
\usepackage{dsfont}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bbold}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{enumitem}
\usepackage{mathrsfs}
\usepackage[margin=0.9 in,headsep=0.3 in]{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{xcolor,colortbl}
\usepackage[format=plain,
labelfont={bf,it},
textfont={it}]{caption}
\usepackage{float}
\usepackage{multirow}

\usepackage[backend=biber, style=ieee]{biblatex}
\addbibresource{Bibliography.bib}

%%%%%%%%%%%NEW PACKAGES I'VE ADDED FOR THE PURPOSE OF THESIS
\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{defn}{Definition}[section] % definition numbers are dependent on theorem numbers

\usepackage{newclude}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
%%Define the two block types and arrow type for our flow diagram
\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=white]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=white]
\tikzstyle{arrow} = [thick,->,>=stealth]
%%%%%%%%%%%END OF NEW PACKAGES I'VE ADDED FOR THE PURPOSE OF THESIS

\newcommand*{\SignatureAndDate}[1]{%
	\par\noindent\makebox[2.5in]{\hrulefill} \hfill\makebox[2.0in]{\hrulefill}%
	\par\noindent\makebox[2.5in][l]{#1}      \hfill\makebox[2.0in][l]{Date}%
}%
\newcolumntype{L}{>{\centering\arraybackslash}m{2cm}}
\newcolumntype{P}{>{\centering\arraybackslash}m{3cm}}
\newcolumntype{Q}{>{\centering\arraybackslash}m{4cm}}
\setlength{\parindent}{0em}

\allowdisplaybreaks

\newcommand{\R}{\mathds{R}}
\newcommand{\Z}{\mathds{Z}}
\newcommand{\Rplus}{\mathds{R}_{> 0}}
\newcommand{\Zplus}{\mathds{Z}_{\geq 0}}
\newcommand{\F}{\mathds{F}}
\newcommand{\N}{\mathds{N}}
\newcommand{\T}{\mathds{T}}
\newcommand{\s}{\mathds{S}}
\newcommand{\C}{\mathds{C}}
\newcommand{\CDFT}{\mathscr{F}_{CD}} %Fourier transform
\newcommand{\ip}[2]{\langle #1, #2\rangle}

\setlength{\parskip}{0.5em}

\usepackage{fancyhdr}
\setlength{\headheight}{15.2pt}
\pagestyle{fancy}
%%\fancyhead[RE]{\thepage}
%%\fancyhead[ER]{\rightmark}
\fancyhead[LE,RO]{\slshape \rightmark}
\fancyhead[LO,RE]{\slshape \leftmark}
\fancyfoot[C]{\thepage}

\makeatletter
\newsavebox\myboxA
\newsavebox\myboxB
\newlength\mylenA

\title{ A-1: Modelling and Control of a Ripstik\textsuperscript{\textregistered} \\MTHE 493 Final Report}
\date{Friday April $7^{\textrm{th}}$, 2017}
\author{\\Andrew Cantanna (10092489) \\Cameron Hudson (10092287)\\
	Benjamin Rudson (10083664)\\}
\begin{document}

	\begin{titlepage}
		\maketitle
		\thispagestyle{empty}		
	\end{titlepage}
	\newpage
	\pagenumbering{Roman} 
	%%\pagestyle{myheadings}
\begin{abstract}
The objective of this project is to develop a comprehensive mathematical model and control system for a RipStik. The system is modeled using principles of Lagrangian mechanics. The Lagrangian is constructed and applied to the Euler-Lagrange equation to develop the unconstrained equations of motion. Nonholonomic constraints are then implemented to the system using numeric integration methods. Linear quadratic regulator control is used to develop a stabilizing control law for an inverted pendulum affixed to a rolling wheel.

A test driven iterative design process is used throughout the course of the project to verify the key mathematical principles governing the model and control system. This framework allows for the studied principals studied to be applied to a variety of complex mechanical systems for use in personal electric transport.

A detailed economic analysis is presented to investigate the scope of the personal electric transport market. Environmental, social and legal considerations pertaining to personal electric transport are also explored.

\end{abstract}	
\clearpage
\newpage
\begin{center}
\textbf{Acknowledgements}
\end{center}
This thesis would not have been possible without the help of many people. 
Firstly, we would like to thank Professor Andrew Lewis for his guidance throughout the project. 
Professor Lewis provided the group with the necessary mathematical background to approach the problem and aided the group in approaching the problem from a pragmatic viewpoint. 
Professor Lewis' wit and candour contributed greatly to the group's enjoyment of the project.
Secondly, we would like to thank Simon Gladstone.
Finally, we would like to thank Benjamin Rudson for taking time out of his rigorous IRONMAN training schedule to attend both the interim and final presentation. 

\vspace*{\fill}
\begin{center}
\textit{When in doubt, close Mathematica and reopen.}
\end{center}
\newpage	
	\renewcommand{\baselinestretch}{0.75}\normalsize
	\tableofcontents
	\renewcommand{\baselinestretch}{1.0}\normalsize

\newpage
\listoffigures
\newpage
\listoftables


\clearpage
\pagenumbering{arabic}
	
	\newpage



	\pagestyle{fancy}
\include{Introduction}
\include{DesignProcess}
\include{Development}
\include{EngineeringConsiderations}
\include{Conclusions}

\newpage
\printbibliography

\end{document}          
